---
title: "ClipTrackAnalysis"
author: "Flora Timney"
date: "2025-04-26"
output: html_document
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```
# Kinematic Analysis of Clips Using trajr
In this script, I load in the tracks from all the dorsal clips and
calculate median and coefficient of variance for speed, the maximum absolute value of acceleration as well as the distance from predator (modeled as the centroid point). I normalise the data, visualize the tracks in space, explore the distributions of these three key kinematic variables within and between species and treatments, and create mixed effect models using them as response variables. 

TL/DR
Botia lohachata species swims faster on average (median) than the Yasuhikotakia modesta species, and fish seem to swim faster when they are closer to the predator; no other significant terms. Speed with a clip varies more (CoV) in the predator treatment than in the control treatment; no other significant terms. Maximum absolute acceleration is also greater when the fish is closer to the predator; no other significant terms. Since the models where mean_dist2pred is significant do not have significant interaction terms of treatment*mean_dist2pred, I think the effect is likely an artefact of the triangular shaped area rather than to do with the predator model.


# I. Set-up 
```{r setup-renv, message = FALSE, warning = FALSE, echo=FALSE}
# renv::init() #initialise renv
renv::snapshot() #update renv
```

```{r setup-wd and clear workspace, message = FALSE, warning = FALSE, echo=FALSE}
setwd(".") #set working directory to where the file is saved
rm(list = ls(all.names = TRUE)) #clear workspace
```

```{r}
source("Functions/packages.r")
source("Functions/extract_metadata.r")
source("Functions/clean_dfs.r")
source("Functions/modelling.r")
source("Functions/calculations.r")
```

Packages:
```{r Install and load packages, message = FALSE, warning = FALSE, echo=FALSE}
packages <- c("trajr", "scales", "tidyverse", "rlang", "bestNormalize", "ggplot2", "ggpubr", "gridExtra", "bbmle", "glmmTMB", "viridis", "emmeans", "fields", "viridisLite", "distributional", "ggdist", "patchwork", "DHARMa", "performance", "gridGraphics", "gt", "webshot2", "magick", "broom.mixed", "conflicted", "dplyr")

check_and_load_packages(packages)
save_package_info()

conflict_prefer("sd", "stats")
conflict_prefer("boxcox", "bestNormalize")
conflict_prefer("chisq.test", "stats")
conflict_prefer("filter", "dplyr")
```

```{r setup-standardise plots, message = FALSE, warning = FALSE, echo=FALSE}
theme_set(theme_minimal()) #standardise the plots

species_colours <- c("Botia_lohachata" = "#E69F00", "Yasuhikotakia_modesta" = "#0072B2")
species_labels <- c(
  "Botia_lohachata" = "Botia lohachata",
  "Yasuhikotakia_modesta" = "Yasuhikotakia modesta"
)
```

```{r}
# Define a wrapper to convert base plot to grob
capture_plot <- function(expr) {
  grid.newpage()
  grid.echo(expr)
  grid.grab()
}
```

## Loading in data

1.  Loading in data and extracting meta data from the clip name

```{r, include=FALSE}
# List all csv files in the directory
track_dir <- "RawData/Tracks"
track_files <- list.files(track_dir, pattern = "_track\\.csv$", full.names = FALSE)

# Optional: remove any non-standard files if needed
track_files <- sort(track_files)
# Remove sub-clips (i.e. parts a, b, c etc)
track_files <- track_files[!grepl("_\\d+[a-z]_", track_files)]

# Read predator data and keep only what we need
pred_df <- read.csv("RawData/pointplot.csv") %>%
  select(TrialID, x_pred, y_pred)

# Define pixel-to-metre scale factor
scale_factor <- 2 / 24.55  # 2cm = 25.55 pixels (measured from circle perimeter/pi = diameter in paddling pool)

# Parse and enrich metadata
meta_data <- tibble(filename = track_files) %>%
  mutate(
    ClipName = str_remove(filename, "_track\\.csv$"),
    Genus = word(ClipName, 1, sep = "_"),
    species = word(ClipName, 2, sep = "_"),
    Treatment = word(ClipName, 3, sep = "_"),
    TrialNum = word(ClipName, 4, sep = "_"),
    ClipNum = word(ClipName, 5, sep = "_"),
    View = word(ClipName, 6, sep = "_"),
    Species = paste(Genus, species, sep = "_"),
    TrialID = paste(Species, Treatment, TrialNum, sep = "_"),
    Group = paste(Species, Treatment, sep = "-")
  ) %>%
  select(filename, ClipName, TrialID, Group, Species, Treatment) %>%
  left_join(pred_df, by = "TrialID") %>%
  mutate(across(c(x_pred, y_pred), ~ .x * scale_factor))


# Build trajectories
csvStruct <- list(x = "x", y = "y", time = "frame", fps = 240)

trjs <- lapply(meta_data$filename, function(fname) {
  df <- read_csv(file.path(track_dir, fname))
  
  raw_trj <- TrajFromCoords(df, xCol = "x", yCol = "y", timeCol = NULL, fps = 240, spatialUnits = "pixels", timeUnits = "frames")
  raw_trj <- TrajScale(raw_trj, scale = scale_factor, unit = "cm")
  
  smooth_trj <- raw_trj
  if (nrow(raw_trj) > 41) {
    smooth_trj <- TrajSmoothSG(raw_trj, p = 3, n = 31)
  }
  
  return(list(raw = raw_trj, smooth = smooth_trj))
})

# Attach ClipNames
names(trjs) <- meta_data$ClipName

# I need Treatment, Species and TrialID to be factors not characters
meta_data$Treatment <- as.factor(meta_data$Treatment)
meta_data$Species <- as.factor(meta_data$Species)
meta_data$TrialID <- as.factor(meta_data$TrialID)

# Extract just the smooth tracks
smoothed_trjs <- lapply(trjs, function(x) x$smooth)
```

## Calculating metrics
2.  Calculating kinematic metrics for each track
```{r}
results <- list()
frame_by_frame_all <- list()

for (clipname in names(smoothed_trjs)) {
  trj <- smoothed_trjs[[clipname]]
  meta_row <- meta_data %>% filter(ClipName == clipname)

  if (nrow(meta_row) == 0 || any(is.na(meta_row$x_pred), is.na(meta_row$y_pred))) next

  pred_pos <- c(meta_row$x_pred, meta_row$y_pred)

  res <- characteriseTrajectory(trj, pred_pos, traj_name = clipname)

  # Add clip name and metadata to frame-by-frame data
  raw <- res$frame_data %>%
    mutate(ClipName = clipname) %>%
    left_join(meta_row, by = "ClipName")

  frame_by_frame_all[[clipname]] <- raw
  results[[clipname]] <- res$summary %>% mutate(ClipName = clipname)
}

# Combine
frame_by_frame_df <- bind_rows(frame_by_frame_all)
summary_df <- bind_rows(results)

# Merge with metadata
meta_stats <- left_join(meta_data, summary_df, by = "ClipName")

# Save optional outputs
write_csv(frame_by_frame_df, "CleanData/frame_by_frame_kinematics.csv")
```


## Normalising data
3.  Checking and normalizing the data distributions (bestNormalize)

I used the bestNormalize() function to work out which transformations were suitable and I ended up deciding that boxcox was best for them all. 
```{r, include=FALSE}
# 1. Median Speed - Box-Cox
trans_median_speed <- boxcox(meta_stats$median_speed)
meta_stats$trans_median_speed <- predict(trans_median_speed)

# 2. Coefficient of Variation of Speed - Log_b(x + a), standardized manually
trans_cv_speed <- boxcox(meta_stats$cv_speed)
meta_stats$trans_cv_speed <- predict(trans_cv_speed)

# 3. Maximum Absolute Acceleration - boxcox
trans_max_acc <- boxcox(meta_stats$max_acc)
meta_stats$trans_max_acc <- predict(trans_max_acc)

write.csv(meta_stats, "CleanData/summary_clip_kinematics.csv", row.names = FALSE)
```

```{r histplots, echo=FALSE, warnings=FALSE, messages=FALSE}
# Create individual ggplot histograms
p1 <- ggplot(meta_stats, aes(x = median_speed)) +
  geom_histogram(fill = "skyblue", color = "white") +
  ggtitle("Median Speed") +
  theme(axis.title.x = element_blank(), axis.text.x = element_blank())

p2 <- ggplot(meta_stats, aes(x = cv_speed)) +
  geom_histogram(fill = "lightgreen", color = "white") +
  ggtitle("Speed CoV") +
  theme(axis.title.x = element_blank(), axis.text.x = element_blank())

p3 <- ggplot(meta_stats, aes(x = max_acc)) +
  geom_histogram(fill = "lightpink", color = "white") +
  ggtitle("Max Abs Acceleration") +
  theme(axis.title.x = element_blank(), axis.text.x = element_blank())

p4 <- ggplot(meta_stats, aes(x = trans_median_speed)) +
  geom_histogram(fill = "skyblue", color = "white") +
  ggtitle("Trans Median Speed") +
  theme(axis.title.x = element_blank(), axis.text.x = element_blank())

p5 <- ggplot(meta_stats, aes(x = trans_cv_speed)) +
  geom_histogram(fill = "lightgreen", color = "white") +
  ggtitle("Trans Speed CoV") +
  theme(axis.title.x = element_blank(), axis.text.x = element_blank())

p6 <- ggplot(meta_stats, aes(x = trans_max_acc)) +
  geom_histogram(fill = "lightpink", color = "white") +
  ggtitle("Trans Max Abs Acceleration") +
  theme(axis.title.x = element_blank(), axis.text.x = element_blank())


# Combine plots in 2 rows, 3 columns
normhist_plot <- (p1 | p2 | p3) / (p4 | p5 | p6) + 
  plot_annotation(
    title = "Normalising the response variables through Box-Cox transformation"
  )

normhist_plot

# Save the combined plot
ggsave("Figures/Kinematics/NormalisingHists.svg", plot = normhist_plot, width = 10, height = 6)
ggsave("Figures/Kinematics/NormalisingHists.png", plot = normhist_plot, width = 10, height = 6)
```

# II. Visualising Tracks

4.  Plotting the tracks in space, subsetted by species-treatment combo
    and coloured by median speed

```{r, results = "hide"}

# Function to convert group names to clean titles
format_group_title <- function(group_string) {
  parts <- unlist(strsplit(group_string, "-"))
  species <- gsub("_", " ", parts[1])
  treatment <- ifelse(parts[2] == "C", "Control", "Predator")
  paste0(species, " - ", treatment)
}

# Create color palette for speed
n_colors <- 100
palette_fn <- colorRampPalette(viridis::viridis(n_colors))
colors <- palette_fn(n_colors)

# Get speed values and color assignments
speed_vals <- meta_stats$trans_median_speed
speed_range <- range(speed_vals, na.rm = TRUE)
speed_indices <- as.integer(cut(speed_vals, breaks = n_colors, include.lowest = TRUE))
speed_colors <- colors[speed_indices]

# Define unique groups
groups <- unique(meta_data$Group)

# Axis limits for consistency
all_x <- unlist(lapply(trjs, function(t) t$smooth$x))
all_y <- unlist(lapply(trjs, function(t) t$smooth$y))
xlim_all <- range(all_x)
ylim_all <- range(all_y)

# Plot function (used for both SVG and PNG)
plot_trajectories <- function(filename, type = c("svg", "png")) {
  if (type == "svg") {
    svg(filename, width = 10, height = 6)
  } else {
    png(filename, width = 1000, height = 600, res = 150)
  }

  par(mfrow = c(2, 2), mar = c(4, 4, 3, 1))

  for (grp in groups) {
    idx <- which(meta_data$Group == grp)
    title <- format_group_title(grp)

    plot(NA, xlim = xlim_all, ylim = rev(ylim_all),
         xlab = "x (m)", ylab = "y (m)", main = title)

    for (i in idx) {
      trj_smooth <- trjs[[i]]$smooth
      lines(trj_smooth, col = speed_colors[i], lwd = 1.5)
    }
  }

  # Add legend
image.plot(legend.only = TRUE, zlim = speed_range, col = colors,
           legend.args = list(text = "Median Speed (m/s)", side = 4, line = 2.5))


  dev.off()
}

# Save both formats
plot_trajectories("Figures/Kinematics/PlotTrajectories.svg", type = "svg")
plot_trajectories("Figures/Kinematics/PlotTrajectories.png", type = "png")

```

This plots helps us visualise lots of things quickly. Firstly, we can see how the clips were distributed across the speceis and treatments and we see that Botia lohachata P is somewhat overrepresented. I am not too worries about the spread because the sample size is so large but if what I captured is truly a 'signalling' behaviour, then it would make sense for the frequency to be highest in the predator treatment and especially in the patterned fish species. Similarly, since the black dots mark origin points, the left to right split is fairly even. Secondly, we can judge by eye how median speed varies with species, treatment and distance to predator (predator model at top); a clear pattern doesn't emerge yet but these factors all captured later in our model.


```{r, eval=FALSE}
library(trajr)
library(ggplot2)
library(gridExtra)
library(svglite)

# Create output directory if needed
out_dir <- "Figures/Kinematics/Trajectories"
if (!dir.exists(out_dir)) dir.create(out_dir, recursive = TRUE)

# Loop through each trajectory
for (i in seq_along(smoothed_trjs)) {
  trj <- smoothed_trjs[[i]]
  trj_name <- names(smoothed_trjs)[i]

  # Data frame with x, y, frame
  df <- trj[, c("x", "y")]
  df$frame <- 1:nrow(df)

  # Calculate velocity and convert to speed (magnitude)
  velocity <- TrajVelocity(trj)
  df$speed <- Mod(velocity)  # make values positive


  # Spatial plot
  p1 <- ggplot(df, aes(x = x, y = y)) +
  geom_path(linewidth = 1) +
  coord_fixed(ratio = 1) +   # fixed 1:1 aspect ratio to keep x,y scales equal
  theme_minimal() +
  theme(
    plot.margin = margin(5, 5, 5, 5),
    panel.spacing = unit(0, "cm")
  ) +
  labs(title = paste("Trajectory:", trj_name),
       x = "x (cm)", y = "y (cm)")

p2 <- ggplot(df, aes(x = frame, y = speed)) +
  geom_line(linewidth = 1, color = "darkblue") +
  theme_minimal() +
  theme(
    plot.margin = margin(5, 5, 5, 5),
    panel.spacing = unit(0, "cm"),
    aspect.ratio = 1  # fix aspect ratio of speed plot to 1:1 too
  ) +
  labs(title = "Speed over time",
       x = "Frame", y = "Speed (cm/s)")

# Combine plots with equal height using patchwork
combined_plot <- p1 | p2 + plot_layout(widths = c(1, 1), guides = "collect") & 
  theme(plot.margin = margin(5, 5, 5, 5))


# File names
png_path <- file.path(out_dir, paste0(trj_name, ".png"))
svg_path <- file.path(out_dir, paste0(trj_name, ".svg"))

ggsave(png_path, plot = combined_plot, width = 10, height = 5, dpi = 300, bg = "white")
ggsave(svg_path, plot = combined_plot, width = 10, height = 5, device = svglite::svglite)

}
```


# III. Exploring Kinematic Metrics

## Plotting distributions
Figure of halfeye plots for all four response variables - distribution of
    kinematic variables across species and treatments

```{r, results = "hide"}
# (a) Median Speed
p1 <- ggplot(meta_stats, aes(y = Treatment, x = trans_median_speed)) +
  stat_halfeye(
    aes(fill = Species, fill_ramp = after_stat(level)),
    .width = c(.66, .95, 1),
    position = "dodgejust"
  ) +
  scale_fill_ramp_discrete(range = c(0.35, 1), na.translate = FALSE) +
  scale_fill_manual(values = species_colours, labels=species_labels) +  # Custom color scale
  labs(title = "Median Speed", 
       y = "Treatment", 
       x = "Normalised Median Speed", fill_ramp = "interval") +
  theme_minimal() +
  theme(legend.position="none",legend.text = element_text(face = "italic"))

# (b) CoV Speed
p2 <- ggplot(meta_stats, aes(y = Treatment, x = trans_cv_speed)) +
  stat_halfeye(
    aes(fill = Species, fill_ramp = after_stat(level)),
    .width = c(.66, .95, 1),
    position = "dodgejust"
  ) +
  scale_fill_ramp_discrete(range = c(0.35, 1), na.translate = FALSE) +
  scale_fill_manual(values = species_colours, labels=species_labels) +  # Custom color scale
  labs(title = "CoV Speed", 
       y = "Treatment", 
       x = "Normalised CoV Speed", fill_ramp = "interval") +
  theme_minimal() +
  theme(legend.position="none", legend.text = element_text(face = "italic"))

# (c) Median Acceleration
p3 <- ggplot(meta_stats, aes(y = Treatment, x = trans_max_acc)) +
  stat_halfeye(
    aes(fill = Species, fill_ramp = after_stat(level)),
    .width = c(.66, .95, 1),
    position = "dodgejust"
  ) +
  scale_fill_ramp_discrete(range = c(0.35, 1), na.translate = FALSE) +
  scale_fill_manual(values = species_colours, labels=species_labels) +  # Custom color scale
  labs(title = "Max Acceleration", 
       xy = "Treatment", 
       x = "Normalised Max Abs Acceleration", fill_ramp = "interval") +
  theme_minimal() +
  theme(legend.position="none", legend.text = element_text(face = "italic"))

# Arrange the plots in a 2x2 grid with common legend
final_plot <- ggarrange(p1, p2, p3, ncol = 3, nrow = 1, 
                        common.legend = TRUE, legend = "bottom",
                        heights = c(1,1), widths = c(1, 1))

# Add a title to the whole figure
halfeye_plot <- annotate_figure(final_plot,
                                         top = text_grob("Distribution of Kinematic Metrics Across Species and Treatments", 
                                                         face = "bold", size = 14))

# Your plotting code
halfeye_plot

# To save to file:
ggsave("Figures/Kinematics/Plots/HalfeyePlot.svg", plot = halfeye_plot, width = 10, height = 5)
ggsave("Figures/Kinematics/Plots/HalfeyePlot.png", plot = halfeye_plot, width = 10, height = 5)

```

Between my four response variables, and is later clear with fitting models, I think my intuition that outright speed would be the key variable is correct. But overall, the central 66% of each distribution (coloured in darker here) does seem to mostly overlap across species and treatments. 


# IV. Frequency of clips

Chi squared goodness-of-fit test to see how my clips are distributed between Species-Treatment Groups. 


```{r}
library(dplyr)
library(tidyr)
library(gt)

# Step 1: Create observed table
observed <- table(meta_stats$Group)

# Step 2: Equal expectations
expected_even <- rep(sum(observed) / length(observed), length(observed))
chisq_test_equal <- chisq.test(observed, p = rep(1/length(observed), length(observed)))
residuals_even <- chisq_test_equal$stdres

chi_table_even <- data.frame(
  Group = names(observed),
  Observed = as.vector(observed),
  Expected = as.vector(expected_even),
  Residuals = round(as.numeric(residuals_even), 2),
  Test = "Equal Proportions"
)

# Step 3: Weighted expectations
group_counts <- meta_stats %>%
  dplyr::select(Group, TrialID) %>%
  dplyr::distinct() %>%
  dplyr::count(Group, name = "UniqueTrialCount") %>%
  dplyr::filter(Group %in% names(observed)) %>%
  dplyr::arrange(factor(Group, levels = names(observed)))

expected_prop <- group_counts$UniqueTrialCount / sum(group_counts$UniqueTrialCount)
expected_weighted <- expected_prop * sum(observed)

chisq_test_weighted <- chisq.test(observed, p = expected_prop)
residuals_weighted <- chisq_test_weighted$stdres

chi_table_weighted <- data.frame(
  Group = names(observed),
  Observed = as.vector(observed),
  Expected = round(expected_weighted, 2),
  Residuals = round(as.numeric(residuals_weighted), 2),
  Test = "Weighted by TrialID"
)

# Step 4: Combine both chi tables
chi_full <- bind_rows(chi_table_even, chi_table_weighted)

# Step 5: Add test-level summary (ChiSq, df, p-value)
summary_long <- data.frame(
  Group = NA,
  Observed = NA,
  Expected = NA,
  Residuals = NA,
  Test = c("Equal Proportions", "Weighted by TrialID"),
  ChiSq = c(as.numeric(chisq_test_equal$statistic), as.numeric(chisq_test_weighted$statistic)),
  df = c(chisq_test_equal$parameter, chisq_test_weighted$parameter),
  p_value = c(chisq_test_equal$p.value, chisq_test_weighted$p.value)
)
summary_reduced <- summary_long %>%
  select(Test, ChiSq, df, p_value)

# Step 6: Join summary back to main table
final_table <- chi_full %>%
  left_join(summary_reduced, by = "Test") %>%
  select(Test, Group, Observed, Expected, Residuals, ChiSq, df, p_value)


# Step 7: Clean and prepare the final table for gt
final_table_cleaned <- final_table %>%
  arrange(Test, Group) %>%
  group_by(Test) %>%
  mutate(
    # Round ChiSq and p_value to 2 decimals before converting to character
    ChiSq = round(ChiSq, 2),
    p_value = round(p_value, 2),
    Test_display = ifelse(row_number() == 1, Test, ""),           # Empty string merges visually
    ChiSq = ifelse(row_number() == 1, as.character(ChiSq), " "),  # Space instead of NA
    df = ifelse(row_number() == 1, as.character(df), " "),
    p_value = ifelse(row_number() == 1, as.character(p_value), " ")
  ) %>%
  ungroup() %>%
  select(-Test) %>%                # Remove original Test column
  rename(Test = Test_display) %>% # Rename Test_display to Test
  select(Test, ChiSq, df, p_value, everything()) # Reorder columns with Test first

# Create and display the gt table
gt_table_comb <- final_table_cleaned %>%
  gt() %>%
  tab_header(title = "Chi-squared Test Results by Group") %>%
  fmt_number(columns = c(Observed, Expected, Residuals), decimals = 3) %>%
  cols_label(
    Test = "Test",
    ChiSq = "Chi-sq",
    df = "df",
    p_value = "p"
  ) %>%
  cols_move_to_start(columns = Test)

# Create separate gt table for Equal Proportions
gt_table_even <- chi_table_even %>%
  gt() %>%
  tab_header(title = "Chi-squared: Equal Proportions") %>%
  fmt_number(columns = c(Observed, Expected, Residuals), decimals = 3) %>%
  cols_label(
    Group = "Group",
    Observed = "Observed",
    Expected = "Expected",
    Residuals = "Std Residual"
  )

# Create separate gt table for Weighted Expectations
gt_table_weighted <- chi_table_weighted %>%
  gt() %>%
  tab_header(title = "Chi-squared: Weighted by TrialID") %>%
  fmt_number(columns = c(Observed, Expected, Residuals), decimals = 3) %>%
  cols_label(
    Group = "Group",
    Observed = "Observed",
    Expected = "Expected",
    Residuals = "Std Residual"
  )

# Save as PNG
gtsave(gt_table_weighted, "Figures/Kinematics/ Models/Frequency/chi_table_weighted.png")
gtsave(gt_table_even, "Figures/Kinematics/ Models/Frequency/chi_table_even.png")
gtsave(gt_table_comb, "Figures/Kinematics/ Models/Frequency/chi_table_comb.png")

# Display the table object
gt_table_comb

```



# V. Modelling: Characteristics of clips

## Establishing models and Saving Outputs

```{r}
source("Functions/modelling.r")
mod1.3 <- glmmTMB(trans_median_speed ~ Species + Treatment + scale(mean_dist2pred) + (1|TrialID), data = meta_stats)

mod2.3 <- glmmTMB(trans_cv_speed ~ Species + Treatment + scale(mean_dist2pred) + (1|TrialID), data = meta_stats)

mod3.3 <- glmmTMB(trans_max_acc ~ Species + Treatment + scale(mean_dist2pred) + (1|TrialID), data = meta_stats)

models <- list(
  "Median Speed" = mod1.3,
  "Speed CoV" = mod2.3,
  "Maximum Absolute Acceleration" = mod3.3
)

# making and saving tidy table outputs of anova summary
imap(models, ~ {
  tbl <- glmmTMB_to_gt(.x, .y)
  out_path <- file.path("Figures","Kinematics", "Models", paste0(.y, "_summary.png"))
  # ensure the directory exists
  dir.create(dirname(out_path), recursive = TRUE, showWarnings = FALSE)
  gtsave(tbl, filename = out_path)
})

# making and saving tidy table outputs of drop1()
iwalk(models, ~ drop1_to_gt(.x, model_name = .y,
                            out_dir = "Figures/Kinematics/Models",
                            file_name = paste0(.y, "_drop1.png")))


# combined summary and drop1 table
iwalk(models, ~ combined_model_summary_gt(
  model      = .x,
  model_name = .y,
  out_dir    = "Figures/Kinematics/Models",
  file_name  = paste0(.y, "_summary_combined.png")
))

# making and saving contvar plots
ContVar_plots <- imap(models, ~ make_contvar_plot(
  model          = .x,
  model_name     = .y,
  data           = meta_stats,
  out_dir        = "Figures/Kinematics/Plots",
  n_points       = 200
))


# making and saving interaction plots
inter_plots <- imap(models, ~ make_inter_plot(
  model          = .x,
  model_name     = .y,
  data           = meta_stats,
  out_dir        = "Figures/Kinematics/Plots",
  width          = 5,
  height         = 5
))

# check model assumptions
iwalk(models, ~ save_model_diagnostics(
  model = .x,
  model_name = .y,
  out_dir = "Figures/Kinematics/Models/Diagnostics"
))
```



##  (a) Median Speed

Patterned B.lohachata is faster than Plain Y. modesta in both treatments - this could be to do with anything about the species (habitat, specific body shape, ecologies etc). Judging from the significant distribution overlap, there doesn't seem to be an effect of treatment.

6a. Applying and simplifying models

1.1 Full model with all two and three way interaction terms
```{r mod1.1, include=FALSE}
mod1.1 <- glmmTMB(trans_median_speed ~ (Species + Treatment + scale(mean_dist2pred))^3 + (1|TrialID), data=meta_stats)
summary(mod1.1)
drop1(mod1.1, test="Chi")
```
The only significant term seems to be Species.

1.2. Model with 3-way interaction removed.
```{r mod1.2, include=FALSE}
mod1.2 <- glmmTMB(trans_median_speed ~ (Species + Treatment + scale(mean_dist2pred))^2 + (1|TrialID), data=meta_stats)
summary(mod1.2)
drop1(mod1.2, test="Chi")
```

Again, Species is the only significant term and none of the interaction terms are significant so, I will remove them.

1.3 All interactions removed
```{r mod1.3, echo=FALSE}
mod1.3 <- glmmTMB(trans_median_speed ~ Species + Treatment + scale(mean_dist2pred) + (1|TrialID), data = meta_stats)
summary(mod1.3)
drop1(mod1.3)
```
Species and distance to predator are the significant terms. I will do no further model simplification. Drop1 test confirms that species and distance to predator are significant terms but not treatment. 

6b. Plots

```{r mod1.1: ContPlot, echo=FALSE}
nseq <- function(x, len=length(x)) seq(min(x, na.rm=TRUE), max(x, na.rm=TRUE), length=len)


pred_data<-data.frame(mean_dist2pred=nseq(meta_stats$mean_dist2pred, nrow(meta_stats)), 
                      Treatment="C", 
                      Species="Botia_lohachata",
                      TrialID=NA)

pred_data_2<-pred_data
pred_data_2$Species="Yasuhikotakia_modesta"

pred_data<-rbind(pred_data, pred_data_2)

new_data<-predict(mod1.3, pred_data, se.fit=TRUE, type="response")
new_data<-data.frame(new_data)
pred_data<-cbind(pred_data, new_data)
colnames(pred_data)<-c("mean_dist2pred", "Treatment", "Species", "TrialID", "trans_median_speed", "s.e.")
pred_data$Upper<-pred_data$trans_median_speed+1.96*pred_data$s.e.
pred_data$Lower<-pred_data$trans_median_speed-1.96*pred_data$s.e.

mod1.3_contvar_plot <- ggplot(data=meta_stats, aes(x=mean_dist2pred, y=trans_median_speed, col=Species)) +
  geom_point() +
  geom_ribbon(data=pred_data, aes(ymin=Lower, ymax=Upper, col=Species, fill=Species), alpha=0.2) +
  geom_line(data=pred_data, aes(x=mean_dist2pred, y=trans_median_speed, col=Species)) +
  scale_fill_manual(values = species_colours) +
  scale_color_manual(values = species_colours) +
  theme_minimal() +
  theme(
    legend.position = c(0.95, 0.95),       # x and y positions: top-right inside plot
    legend.justification = c("right", "top"),  # Anchor legend by its top-right corner
    legend.background = element_rect(fill = alpha("white", 0.7), color = NA),  # Semi-transparent bg
    legend.key = element_rect(fill = "transparent", color = NA)
  )

mod1.3_contvar_plot
ggsave("Figures/Plots/Kinematics/Mod1.3_ContVar_plot.svg", plot = mod1.3_contvar_plot, width = 5, height = 5)

```

The significant term in the model is the Species. Botia lohachata is faster than Yasuhikotakia modesta.

7b. Plotting model results with interactions

```{r mod1.3 InterPlot, echo=FALSE, warning=FALSE, message=FALSE}
# Run ANOVA

mod1.3_inter_plot <- ggline(meta_stats, x = "Treatment", y = "trans_median_speed", color = "Species",
       add = c("mean_se", "dotplot"),
       palette = species_colours)

mod1.3_inter_plot
ggsave("Figures/Kinematics/Plots/Mod1.3_Inter_plot.svg", plot = mod1.3_inter_plot, width = 5, height = 5)

```

These interaction plots tell the same story as the box and whisker and halfeye plots from earlier. Botia lohachata is faster overall and slows down notably when in the predator treatment whereas Yasuhikotakia modesta is slower overall and barely changes speed between treatments




--------------------------------------------------------------------------------
##  (b) CoV speed

2.1. Full model with all two and three way interaction terms
```{r mod2.1, include=FALSE}
mod2.1 <- glmmTMB(trans_cv_speed ~ (Species + Treatment + scale(mean_dist2pred))^3 + (1|TrialID), data=meta_stats)
summary(mod2.1)
drop1(mod2.1, test="Chi")
```

No significant terms. Removing 3-way interaction

2.2. Near Full model with all two way interaction terms
```{r mod2.2, include=FALSE}
mod2.2 <- glmmTMB(trans_cv_speed ~ (Species + Treatment + scale(mean_dist2pred))^2 + (1|TrialID), data=meta_stats)
summary(mod2.2)
drop1(mod2.2, test="Chi")
```

No significant terms. Removing 2-way interactions

2.3. Simplified model with all no interaction terms
```{r mod2.3, echo=FALSE}
mod2.3 <- glmmTMB(trans_cv_speed ~ Species + Treatment + scale(mean_dist2pred) + (1|TrialID), data=meta_stats)
summary(mod2.3)
drop1(mod2.3, test="Chi")
```

Now, Treatment and distance to predator are significant terms. Species almost but not quite. 



```{r mod2.3 ContPlot, echo=FALSE}
nseq <- function(x, len=length(x)) seq(min(x, na.rm=TRUE), max(x, na.rm=TRUE), length=len)


pred_data<-data.frame(mean_dist2pred=nseq(meta_stats$mean_dist2pred, nrow(meta_stats)), 
                      Treatment="C", 
                      Species="Botia_lohachata",
                      TrialID=NA)

pred_data_2<-pred_data
pred_data_2$Species="Yasuhikotakia_modesta"

pred_data<-rbind(pred_data, pred_data_2)

new_data<-predict(mod2.3, pred_data, se.fit=TRUE, type="response")
new_data<-data.frame(new_data)
pred_data<-cbind(pred_data, new_data)
colnames(pred_data)<-c("mean_dist2pred", "Treatment", "Species", "TrialID", "trans_cv_speed", "s.e.")
pred_data$Upper<-pred_data$trans_cv_speed+1.96*pred_data$s.e.
pred_data$Lower<-pred_data$trans_cv_speed-1.96*pred_data$s.e.

mod2.3_contvar_plot <- ggplot(data = meta_stats, aes(x = mean_dist2pred, y = trans_cv_speed, col = Species)) +
  geom_point() +
  geom_ribbon(data = pred_data, aes(x = mean_dist2pred, ymin = Lower, ymax = Upper, fill = Species), alpha = 0.2, color = NA) +  # no color for ribbon edges
  geom_line(data = pred_data, aes(x = mean_dist2pred, y = trans_cv_speed, col = Species)) +
  scale_fill_manual(values = species_colours) +
  scale_color_manual(values = species_colours) +
  theme_minimal() +
  theme(
    legend.position = c(0.95, 0.95),       # inside top-right
    legend.justification = c("right", "top"),
    legend.background = element_rect(fill = alpha("white", 0.7), color = NA),
    legend.key = element_rect(fill = "transparent", color = NA)
  )

mod2.3_contvar_plot
ggsave("Figures/Kinematics/Plots/Mod2.3_ContVar_plot.svg", plot = mod2.3_contvar_plot, width = 5, height = 5)

```

```{r mod2.3 InterPlot, echo=FALSE, message=FALSE, warning=FALSE}
mod2.3_inter_plot <- ggline(meta_stats, x = "Treatment", y = "trans_cv_speed", color = "Species",
       add = c("mean_se", "dotplot"),
       palette = species_colours)

mod2.3_inter_plot
ggsave("Figures/Kinematics/Plots/Mod2.3_Inter_plot.svg", plot = mod2.3_inter_plot, width = 5, height = 5)
```

--------------------------------------------------------------------------------
## (c) Max absolute acceleration

3.1. Full model with all two and three way interaction terms
```{r mod3.1, include=FALSE}
mod3.1 <- glmmTMB(trans_max_acc ~ (Species + Treatment + scale(mean_dist2pred))^3 + (1|TrialID), data=meta_stats)
summary(mod3.1)
drop1(mod3.1, test="Chi")
```

No significant terms. Removing 3-way interaction

3.2. Near Full model with all two way interaction terms
```{r mod3.2, include=FALSE}
mod3.2 <- glmmTMB(trans_max_acc ~ (Species + Treatment + scale(mean_dist2pred))^2 + (1|TrialID), data=meta_stats)
summary(mod3.2)
drop1(mod3.2, test="Chi")
```

Minimum distance to predator is significant. No significant interaction terms. Removing 2-way interactions

3.3. Simplified model with all no interaction terms
```{r mod3.3, echo=FALSE}
mod3.3 <- glmmTMB(trans_max_acc ~ Species + Treatment + scale(mean_dist2pred) + (1|TrialID), data=meta_stats)
summary(mod3.3)
drop1(mod3.3, test="Chi")
```

Distance to predator is the only significant term. 

6c. Plotting mean speed against distance to predator

```{r mod3.3 ContPlot, echo=FALSE}
nseq <- function(x, len=length(x)) seq(min(x, na.rm=TRUE), max(x, na.rm=TRUE), length=len)


pred_data<-data.frame(mean_dist2pred=nseq(meta_stats$mean_dist2pred, nrow(meta_stats)), 
                      Treatment="C", 
                      Species="Botia_lohachata",
                      TrialID=NA)

pred_data_2<-pred_data
pred_data_2$Species="Yasuhikotakia_modesta"

pred_data<-rbind(pred_data, pred_data_2)

new_data<-predict(mod3.3, pred_data, se.fit=TRUE, type="response")
new_data<-data.frame(new_data)
pred_data<-cbind(pred_data, new_data)
colnames(pred_data)<-c("mean_dist2pred", "Treatment", "Species", "TrialID", "trans_max_acc", "s.e.")
pred_data$Upper<-pred_data$trans_max_acc+1.96*pred_data$s.e.
pred_data$Lower<-pred_data$trans_max_acc-1.96*pred_data$s.e.

mod3.3_contvar_plot <- ggplot(data = meta_stats, aes(x = mean_dist2pred, y = trans_max_acc, col = Species)) +
  geom_point() +
  geom_ribbon(
    data = pred_data,
    aes(x = mean_dist2pred, ymin = Lower, ymax = Upper, fill = Species),
    alpha = 0.2,
    color = NA  # removes ribbon borders
  ) +
  geom_line(
    data = pred_data,
    aes(x = mean_dist2pred, y = trans_max_acc, col = Species)
  ) +
  scale_fill_manual(values = species_colours) +
  scale_color_manual(values = species_colours) +
  theme_minimal() +
  theme(
    legend.position = c(0.95, 0.95),  # top-right inside plot
    legend.justification = c("right", "top"),
    legend.background = element_rect(fill = alpha("white", 0.7), color = NA),
    legend.key = element_rect(fill = "transparent", color = NA)
  )

mod3.3_contvar_plot
ggsave("Figures/Kinematics/Plots/Mod3.3_ContVar_plot.svg", plot = mod3.3_contvar_plot, width = 5, height = 5)

```

```{r mod3.3 Interplot, echo=FALSe, message=FALSe, warning=FALSE}
mod3.3_inter_plot <- ggline(meta_stats, x = "Treatment", y = "trans_max_acc", color = "Species",
       add = c("mean_se", "dotplot"),
       palette = c("#E69F00", "#0072B2"))

mod3.3_inter_plot
ggsave("Figures/Kinematics/Plots/Mod3.3_Inter_plot.svg", plot = mod3.3_inter_plot, width = 5, height = 5)

```


-------------------------------------------------------------------------------

## Things to fix
I want to save smoothed_trjs
```{r}
# library(trajr)
# write.trajr(smoothed_trjs, file = "smoothed_trjs.csv")
```

I also want to print out the model summary results neatly. 
```{r eval=FALSE}
library(glmmTMB)
library(broom.mixed)  # for tidy.glmmTMB
library(dplyr)
library(pander)
library(grid)


# Tidy model coefficients (fixed effects)
tidy_mod <- tidy(mod1.3, effects = "fixed")

# drop1 results as data.frame
drop1_df <- as.data.frame(drop1(mod1.3, test = "Chi"))
drop1_df$term <- rownames(drop1_df)
rownames(drop1_df) <- NULL
tidy_mod$term <- c("Intercept", "Species", "Treatment", "scale(mean_dist2pred)")

# Merge the two tables by term
merged_df <- tidy_mod %>%
  full_join(drop1_df, by = "term") %>%
  filter(term != "<none>") %>%  # â† Remove the intercept row
  select(term, estimate, std.error, statistic, p.value, `Pr(>Chi)`)

pander(merged_df, caption = "Model 1.3: Coefficients and drop1 Results")

# Round numeric values for display
merged_print <- merged_df %>%
  mutate(across(where(is.numeric), ~ round(., 3)))

# Convert to table grob
table_grob <- tableGrob(merged_print)

# Save as SVG
svg("Figures/Kinematics/Models/model1.3_results.svg", width = 8, height = 5)
grid.draw(table_grob)
dev.off()
```


# T-test of fish height
```{r}
heightdata <- read_csv("RawData/FishHeight.csv")
head(heightdata)

t_result <- t.test(Height ~ Species, data = heightdata)
print(t_result)
```

